\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-05-03 forwards-and-backwards.tex
.
one bib file
final final edits, sending to co-authors

\comment{TODO: signpost where proofs can be found}

@@

We shall consider three different classes of scaling limits for our model.
The first yields a class of (generalised) superprocesses in which coefficients 
governing both the spatial motion and the branching components of the process can depend
on local population density; the second is a corresponding class of deterministic
non-local differential equations; and the third are classical PDEs.
We measure local population density around a point by convolving with
a smooth kernel $\rho(\cdot)$, which may differ
for the two stages of reproduction. 
When the limiting population process is deterministic,
it is a (weak) solution of an equation of the form
\begin{equation}
	\label{general deterministic limit}
        \partial_t \varphi_t(x)
        =
        r\left(x, \varphi_t \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \varphi_t \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \varphi_t \right)
        ,
\end{equation}
where
$\varphi_t(x)$ can be thought of as the population density at $x$
(although the limit may be a measure without a density), and
$\DG^*$ is (the adjoint of) a strictly uniformly elliptic second order differential operator, typically the Laplacian.
The dependence of each of the terms $r$, $\gamma$, and $F$ on $\varphi$ is only through the local density at $x$,
e.g., $F(x, \varphi) = F(x, \smooth{} \varphi(x))$.
We shall be more specific about the parameters below.

@@

Specialising the results of 
\citet{kurtz/xiong:1999} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

To make this precise,
define the spatial projection maps
$\kappa^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\kappa^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\kappa: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\kappa(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a $\sigma$-field $\mathcal{F}$,
we say that \emph{$\lp^N$ is conditionally uniform given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle H_f^N, \kappa^N(\lp) \rangle} ,
\end{align}
where
$$ H_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of 
$\kappa^N(\lp)$.
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \kappa(\lp)(dx) \rangle} .
\end{align}
In other words, $\lp$ is conditionally Poisson with Cox measure $\kappa(\lp) \times \lambda$,
where $\lambda$ is Lesbegue measure.

@@

to obtain 

@@

Consulting~\eqref{eqn:prelimit_martingale_variation},
the quadratic variation of $\widehat{M}^N(f)$ satisfies
$ \IE[[\widehat{M}^N_f]_T] \le C (\theta/N) \int_0^T \langle 1, \eta_s \rangle ds \le C' \theta/N$
for some constants $C$ and $C'$.
Now, by the Burkholder-Davis-Gundy inequality,
$\IE[\sup_{0 \le u \le T} |\widehat{M}^N(f)_u|^2] \le C'' \IE[[\widehat{M}^N(f)]_u]$,
and so using Markov's inequality,
\begin{equation}
\label{martingale term to zero}
    \limsup_{N\to\infty}\IP\Big[
        2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|
        >\frac{\nu}{3}
    \Big]
    \leq
    \limsup_{N\to\infty} \frac{36}{\nu^2} C'' \IE\big[[\widehat{M}^N(f)]_T\big]
    \leq
    \limsup_{N\to\infty}\frac{36}{\nu^2}\frac{C' C''\theta}{N\epsilon^d} 
    =0. 
\end{equation}

@@

\begin{lemma}
    \label{Lemma:BoundPsiHS2}
    In the notation of Lemma~\ref{PsiBoundHS},
    %and with $A$ as in Lemma~\ref{lem:poisson_ld},
	\begin{equation}
\label{heat kernel estimate}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        %\\
\le
        \frac{C}{(\epsilon^2 \theta)^{1/2}}
        p_{6(\epsilon^2+t)}(x, y)
        +
	    \frac{C}{(\epsilon^2 + t)^{d/2}}
	    \exp(- \epsilon^2 \theta / 2)
        %+
        %\frac{C}{(\epsilon^2 + t)^{d/2}}
        %\IP\left\{
        %    \tau(t) > A(\epsilon^2 + t)
        %\right\} 
	    .
	\end{equation}
\end{lemma}

@@

= |\langle p_{\epsilon^2}(y,z)-p_{\epsilon^2}(w,z),\eta_s(dz) \rangle | 

@@

&\qquad

@@

\big\langle |p_{\epsilon^2+s}(y,z)-p_{\epsilon^2+s}(w,z)|, 
\eta_0(dz)\big\rangle 

@@

\big\langle (p_{s-r+\epsilon^2}(y,z)-p_{s-r+\epsilon^2}(w,z)|)
\cdot |F(\rho_\epsilon*\eta_r(z)|, 

@@

	+ \big\langle\frac{C}{(\theta\epsilon^2)^{1/2}} \Big(
	p_{6(\epsilon^2+s)}(y,z)+p_{6(\epsilon^2+s)}(w,z)\Big)
	+\widehat{\epsilon}(\delta,\epsilon,\theta)
	,\eta_0(dz)\rangle

@@

\cdot |F(\rho_\epsilon*\eta_r(z)|, 

@@

%\cdot |F(\rho_\epsilon*\eta_r(z)|, 

@@

\leq \Big\langle \frac{\|y-w\|}{\sqrt{s+\epsilon^2}}

@@

	+ \big\langle \frac{C}{(\theta\epsilon^2)^{1/2}}\Big( 

@@

	\widehat{\epsilon}(\delta,\epsilon,\theta) , \eta_0(dz)
\big\rangle

@@

We now turn to the identification of the limit points of the sequence 
$\big\{\big(\rho_\epsilon*\eta_t^N(x)dx\big)_{t\geq 0}\big\}_{N\geq 1}$.
We would like to show that any limit point solves~\eqref{target equation in local convergence}
in the limit, i.e.,
\begin{equation}\label{eqn:target PDE weak form}
    \langle f(x), \varphi(t,x)dx \rangle
    =
    \int_0^t
        \Big\langle
        \frac{1}{2}\Delta f(x) + f(x) F(\varphi(s,x))
        ,
        \varphi(s,x) dx
        \Big\rangle
    ds .
\end{equation}
Since 
$\langle f, \rho_\epsilon*\eta_t^N(x)dx\rangle=
\langle \rho_\epsilon*f(x), \eta_t^N(dx)\rangle$,
and the limit is deterministic,
this will follow if we can show that each the terms in the semimartingale
decomposition~(\ref{eqn:eta_martingale}),
with the test function $f$ replaced by $\rho_\epsilon*f(\cdot)$,
converges to the corresponding term in~\eqref{eqn:target PDE weak form}.

@@

We now turn to the contribution arising from the martingale terms 
in~(\ref{difference in rhos}):
\[
\IE\Big[
\int_\delta^t\Big\langle\int 
\big(|M_s(y)|+|M_s(w)|\big)\big(\rho_\epsilon*\eta_s(y)^k+
\rho_\epsilon*\eta_s(w)^k\big)p_{\epsilon^2}(w,y)dy, \eta_s(dw)\Big\rangle ds
\Big] .
\]
Since $\psi_{t-s}^{\epsilon,x}(y)=\IE[p_{T(t-s)+\epsilon^2}(x,y)]$, 
rearranging~(\ref{expn rhoepsilon})
we see that we can pull a convolution with $p_{\epsilon^2/2}$ out of
our expression for $M_s(w)$ and so all the manipulations 
that we used to control terms above with still be valid.
To deal with the two terms in the product involving $|M_s(y)|$, we write 
the first as $\int |M_s(y)|\rho_\epsilon*\eta_s(y)^{k+1}dy$ and then use
H\"older's inequality, Lemma~\ref{bounds on moments}, and the fact
that $\IE\big[|M_s(y)|^2\big]$ is ${\mathcal O}\big(\theta/(N\epsilon^d)\big)$
to see that the contribution from this term tends to zero in the limit.
For the second, we use the idea of the proof of 
Corollary~\ref{alternative form moment bounds} to reduce to a form to which
we can apply H\"older's inequality.

@@

In addition to tightness of the measure-values processes $\lp^N$,
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction
of Section~\ref{sec: individual lines of descent} holds.
	\comment{Is this paragraph really a separate remark that we want to make?}
    \comment{PETER: maybe? Figuring out how to make it more concrete/rigorous seems daunting, though -
        I'm inclined to remove it.}

@@

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported
    (following, e.g., \citet{etheridge/kurtz:2019}, Condition 2.1).
    % PLR: NEW DETAILS
    Several places below we require a dense subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, 
    i.e., for any $\lp_1$ and $\lp_2$ there is an $f$
    with $\langle f, \lp_1\rangle \neq \langle f, \lp_2\rangle$,
    and a $g$ such that $\langle g, \lp_1\rangle \neq 0$.
	Therefore,
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    % END NEW DETAILS
    Topologized in this way, the space $\lpmeasures$ completely metrizable,
    and we may choose a countable set of nonnegative $f_k$,
    each supported on $\IR^d \times [0,u_k]$ for some $u_k < \infty$,
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$.
    (To see this, use Theorem A.2.3 of~\citet{kallenberg1997foundations}
    and the first argument made in the proof of Lemma~\ref{lem:product_tightness}.)
    Below, Lemma~\ref{lem:compact containment lookdown} proves exactly this,
    and therefore compact containment.
    Here we have compactified $\IR^d$ for convenience
    (since it turned out to be straightforward to show that mass does not escape to infinity in space);
    however, we need to use the vague topology ``in the level direction''
    because \emph{levels} may escape to infinity in finite time in the limit.

>>>>

@@

We shall consider three different classes of scaling limits for our model.
The first yields a class of (generalised) superprocesses in which coefficients 
governing both the spatial motion and the branching components of the process can depend
on local population density; the second is a corresponding class of deterministic
non-local differential equations; and the third are classical PDEs.
We measure local population density around a point by convolving with
a smooth kernel $\rho(\cdot)$, which may differ
for the two stages of reproduction. 
When the limiting population process is deterministic,
it is a (weak) solution of an equation of the form
\begin{equation}
	\label{general deterministic limit}
        \partial_t \varphi_t(x)
        =
        r\left(x, \varphi_t \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big(\cdot, \varphi_t \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \varphi_t \right)
        ,
\end{equation}
where
$\varphi_t(x)$ can be thought of as the population density at $x$
(although the limit may be a measure without a density), and
$\DG^*$ is (the adjoint of) a strictly uniformly elliptic second order differential operator, typically the Laplacian.
The dependence of each of the terms $r$, $\gamma$, and $F$ on $\varphi$ is only through the local density at $x$,
e.g., $F(x, \varphi) = F(x, \smooth{} \varphi(x))$.
We shall be more specific about the parameters below.

@@

Specialising the results of 
\citet{kurtz/xiong:1999} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \citet{kurtz/xiong:1999} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

To make this precise,
define the spatial projection maps
$\kappa^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\kappa^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\kappa: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\kappa(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a $\sigma$-field $\mathcal{F}$,
we say that \emph{$\lp^N$ is conditionally uniform given $\mathcal{F}$}
if $\kappa^N(\lp)$ is $\mathcal{F}$-measurable and 
for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle H_f^N, \kappa^N(\lp) \rangle} ,
\end{align}
where
$$ H_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of 
$\kappa^N(\lp)$.
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if $\kappa(\lp)$ is $\mathcal{F}$-measurable and 
for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \kappa(\lp)(dx) \rangle} .
\end{align}
In other words, $\lp$ is conditionally Poisson with Cox measure $\kappa(\lp) \times \lambda$,
where $\lambda$ is Lesbegue measure.

@@

to yield 

@@

Consulting~\eqref{eqn:prelimit_martingale_variation},
and using the approach of Lemma~\ref{lem:limit_mgale},
the quadratic variation of $\widehat{M}^N(f)$ satisfies
$ \IE[[\widehat{M}^N_f]_T] \le C (\theta/N) \int_0^T \IE[\langle 1, \eta_s \rangle] ds \le C' \theta/N$
for some constants $C$ and $C'$.
Now, by the Burkholder-Davis-Gundy inequality,
$\IE[\sup_{0 \le u \le T} |\widehat{M}^N(f)_u|^2] \le C'' \IE[[\widehat{M}^N(f)]_u]$,
and so using Markov's inequality,
\begin{equation}
\label{martingale term to zero}
    \limsup_{N\to\infty}\IP\Big[
        2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|
        >\frac{\nu}{3}
    \Big]
    \leq
    \limsup_{N\to\infty} \frac{36}{\nu^2} C'' \IE\big[[\widehat{M}^N(f)]_T\big]
    \leq
    \limsup_{N\to\infty}\frac{36}{\nu^2}\frac{C' C''\theta}{N\epsilon^d} 
    =0. 
\end{equation}

@@

\begin{lemma}
    \label{Lemma:BoundPsiHS2}
    In the notation of Lemma~\ref{PsiBoundHS},
    there exists a $C < \infty$ such that
    %and with $A$ as in Lemma~\ref{lem:poisson_ld},
	\begin{equation}
\label{heat kernel estimate}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        %\\
\le
        \frac{C}{(\epsilon^2 \theta)^{1/2}}
        p_{6(\epsilon^2+t)}(x, y)
        +
	    \frac{C}{(\epsilon^2 + t)^{d/2}}
	    \exp(- \epsilon^2 \theta / 2)
        %+
        %\frac{C}{(\epsilon^2 + t)^{d/2}}
        %\IP\left\{
        %    \tau(t) > A(\epsilon^2 + t)
        %\right\} 
	    .
	\end{equation}
\end{lemma}

@@

    |\langle p_{\epsilon^2}(y,z)-p_{\epsilon^2}(w,z),\eta_s(dz) \rangle | 

@@

@@

\big\langle
    |p_{\epsilon^2+s}(y,z)-p_{\epsilon^2+s}(w,z)|, 
\eta_0(dz) \big\rangle 

@@

    \big\langle |p_{s-r+\epsilon^2}(y,z)-p_{s-r+\epsilon^2}(w,z)||
    |F(\rho_\epsilon*\eta_r(z)|, 

@@

    \frac{C}{(\theta\epsilon^2)^{1/2}}
    \Big(p_{6(\epsilon^2+s)}(y,z)+p_{6(\epsilon^2+s)}(w,z)\Big)
	+ C\widehat{\epsilon}(\delta,\epsilon,\theta)
,\eta_0(dz) \Big\rangle

@@

|F(\rho_\epsilon*\eta_r(z)|, 

@@

%|F(\rho_\epsilon*\eta_r(z)|, 

@@

\Big\langle \frac{\|y-w\|}{\sqrt{s+\epsilon^2}}

@@

\Big\langle \frac{C}{(\theta\epsilon^2)^{1/2}}\Big( 

@@

	C \widehat{\epsilon}(\delta,\epsilon,\theta) , \eta_0(dz)
\Big\rangle

@@

We now turn to the identification of the limit points of the sequence 
$\big\{\big(\rho_\epsilon*\eta_t^N(x)dx\big)_{t\geq 0}\big\}_{N\geq 1}$.
We would like to show that any limit point solves~\eqref{target equation in local convergence}
in the limit, i.e.,
\begin{equation}\label{eqn:target PDE weak form}
    \langle f(x), \varphi(t,x)dx \rangle
    =
    \int_0^t
        \Big\langle
        \frac{1}{2}\Delta f(x) + f(x) F(\varphi(s,x))
        ,
        \varphi(s,x) dx
        \Big\rangle
    ds .
\end{equation}
Since 
$\langle f, \rho_\epsilon*\eta_t^N(x)dx\rangle=
\langle \rho_\epsilon*f(x), \eta_t^N(dx)\rangle$,
and the limit is deterministic,
this will follow if we can show that each of the terms in the semimartingale
decomposition~(\ref{eqn:eta_martingale}),
with the test function $f$ replaced by $\rho_\epsilon*f(\cdot)$,
converges to the corresponding term in~\eqref{eqn:target PDE weak form}.

@@

We now turn to the contribution arising from the martingale terms 
in~(\ref{difference in rhos}):
\[
\IE\Big[
\int_\delta^t\Big\langle\int 
\big(|M_s(y)|+|M_s(w)|\big)\big(\rho_\epsilon*\eta_s(y)^k+
\rho_\epsilon*\eta_s(w)^k\big)p_{\epsilon^2}(w,y)dy, \eta_s(dw)\Big\rangle ds
\Big] .
\]
Since $\psi_{t-s}^{\epsilon,x}(y)=\IE[p_{T(t-s)+\epsilon^2}(x,y)]$, 
rearranging~(\ref{expn rhoepsilon})
we see that we can pull a convolution with $p_{\epsilon^2/2}$ out of
our expressions for $M_s(y)$ and $M_s(w)$ and so all the manipulations 
that we used to control terms above with still be valid.
To deal with the two terms in the product involving $|M_s(y)|$, we write 
the first as $\int |M_s(y)|\rho_\epsilon*\eta_s(y)^{k+1}dy$ and then use
H\"older's inequality, Lemma~\ref{bounds on moments}, and the fact
that $\IE\big[|M_s(y)|^2\big]$ is ${\mathcal O}\big(\theta/(N\epsilon^d)\big)$
to see that the contribution from this term tends to zero in the limit.
For the second, we use the idea of the proof of 
Corollary~\ref{alternative form moment bounds} to reduce to a form to which
we can apply H\"older's inequality.

@@

\begin{remark}
In addition to tightness of the measure-values processes $\lp^N$,
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction
of Section~\ref{sec: individual lines of descent} holds.
\end{remark}

@@

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported
    (following, e.g., \citet{etheridge/kurtz:2019}, Condition 2.1).
    % PLR: NEW DETAILS
    In several places below we require a dense subset
    of $C_b(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $C_b(\lpmeasures)$,
    but they do separate points and vanish nowhere, 
    i.e., for any $\lp_1$ and $\lp_2$ there is an $f$
    with $\langle f, \lp_1\rangle \neq \langle f, \lp_2\rangle$,
    and a $g$ such that $\langle g, \lp_1\rangle \neq 0$.
	Therefore,
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $C_b(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    % END NEW DETAILS
    Topologized in this way, the space $\lpmeasures$ is completely metrizable,
    and we may choose a countable set of nonnegative $f_k$,
    each supported on $\IR^d \times [0,u_k]$ for some $u_k < \infty$,
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$.
    (To see this, use Theorem A.2.3 of~\citet{kallenberg1997foundations}
    and the first argument made in the proof of Lemma~\ref{lem:product_tightness}.)
    Below, Lemma~\ref{lem:compact containment lookdown} proves exactly this,
    and therefore compact containment.
    Here we have compactified $\IR^d$ for convenience
    (since it turned out to be straightforward to show that mass does not escape to infinity in space);
    however, we need to use the vague topology ``in the level direction''
    because \emph{levels} may escape to infinity in finite time in the limit.

    In order to apply \citet{ethier/kurtz:1986} Theorem~3.9.1
    we require that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes, for all $F$ in a subset
    of $C_b(\lpmeasures)$ %, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    % The functions $\lp \mapsto \langle f, \lp \rangle$
    % do not form not a dense subset of $C_b(\lpmeasures)$,
    % but they do separate points and vanish nowhere, 
    % i.e., for any $\lp_1$ and $\lp_2$ there is an $f$
    % with $\langle f, \lp_1\rangle \neq \langle f, \lp_2\rangle$,
    % and a $g$ such that $\langle g, \lp_1\rangle \neq 0$.
	% Therefore,
    % by the Stone-Weierstrass theorem, the algebra they generate is dense
    % % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % % Since $C_0(\lpmeasures)$ is dense
    % in $C_b(\lpmeasures)$ (with respect to uniform convergence on compact subsets).  So, 
    Since as above the algebra generated by the functions $\lp \mapsto \langle f \lp \rangle$
    is dense in $C_b(\lpmeasures)$,
    it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.
    The fact that martingale properties are preserved under passage to the limit
    is straightforward, and can be proved in a way analogous to Lemma~\ref{lem:limit_mgale};
    we omit the proof.
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    By Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    Thus any limit point of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}
